---
   title: "hw1"  
   author: "Adam Pearce"  
   output:  
     html_document:  
       keep_md: true  
---
```{r setup, include = FALSE, cache = FALSE}
knitr::opts_chunk$set(error = TRUE)
```

```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(plyr)
library(dplyr)
library(ggplot2)
library(GGally)
```

Load data
```{r, echo=FALSE}
mobility <- read.csv("mobility.csv", header=TRUE)
```

Mobility map 

```{r, echo=FALSE}
ggplot(mobility  %>% filter(Longitude > -130), aes(x=Longitude, y=Latitude, color=Mobility)) + geom_point(size=5)
```

Mobility is lowest in the South and on the coasts; it is highest in the midwest.

Population v Mobility
```{r}
ggplot(mobility, aes(x=Population, y=Mobility)) + scale_x_log10() + geom_smooth(lwd = 3, se = FALSE, method = "lm") + geom_point(alpha=1/3, size=3)

lm(formula = Mobility ~ log10(Population), data=mobility

```

Areas with higher populations tend towards lower mobility. This relationship doesn't look linear. 

```{r}
mobility %>% 
  select(Mobility, Income, Seg_racial, Share01, School_spending, Violent_crime, Commute) %>%
  ggpairs(axisLabels="none") + 
  theme_bw()
```

None of these correlations looks too odd, expect perhaps mean income being positivly coorelated with mobility. Segreation, violent crime and the 1%'s share of income are negitivly correlated while commute time and schools are positivily correlate with mobility.

```{r}
fit <- lm(formula = Mobility ~ log10(Population) + Income + Seg_racial + School_spending + Share01, data=mobility)
summary(fit)
```

None of the signs have changed.

```{r}
 p <- mobility  %>% filter(Name == 'Pittsburgh')

predict.lm(object = fit)
```
Actual mobility is .094, predicted is 0.12487204





