---
   title: "Functions"  
   author: "Adam Pearce"  
   output:  
     html_document:  
       keep_md: true  
---

```{r, echo=FALSE, message=FALSE, warning=FALSE}
library(plyr)
library(dplyr)
library(ggplot2)
library(robustbase)

gdURL  <-  "http://tiny.cc/gapminder"
gDat  <- read.delim(file = gdURL)

gtbl <- tbl_df(gDat)
```

Fit a regression using ordinary least squares and a robust technique.

```{r message=FALSE, warning=FALSE}
j_coefs <- ddply(gDat, ~ country + continent, function(gDat, offset = 1952) {
  ols <- lm(lifeExp ~ I(year - offset), gDat)
  rbs <- lmrob(lifeExp ~ I(year - offset), gDat, setting = "KS2011")
  ols_coef <- setNames(coef(ols), c("o_int", "o_slope"))
  rbs_coef <- setNames(coef(rbs), c("r_int", "r_slope"))
  dif_coef <- setNames(c(ols_coef[1] - rbs_coef[1], ols_coef[2] - rbs_coef[2]),
                       c("d_int", "d_slope"))
  c(ols_coef, rbs_coef, dif_coef)
})

j_coefs
```

Plot the difference in estimated parameters under the two approaches
```{r}
p  <- ggplot(j_coefs, aes(x = d_slope, y = d_int, label = country)) + geom_point()
p + geom_text(hjust=0, vjust=0, size=3)
```

China and Zimbabwe are the biggest outliers.

```{r}
j_coefs  %>% filter(country == 'China' | country == 'Zimbabwe')
```

The robust regression throws out some of the post WWII and great leap forward years
```{r}
gDat  %>% filter(country == 'China')

ggplot(gDat  %>% subset(country == 'China'), aes(x = year, y=lifeExp)) + geom_point()
```


The opposite happens with Zimbabwe - the  linear increases in life expectency before the 90s results in the robust regression giving a very low weight to the subsequent 'outlier' years. 
```{r}
gDat  %>% filter(country == 'Zimbabwe')

ggplot(gDat  %>% subset(country == 'Zimbabwe'), aes(x = year, y=lifeExp)) + geom_point()
```



Color dots by weight of robust regression?


Divergence between increases in lifeExp and gdpPerCap?




















