---
title: "hw"
output:  
  html_document:  
    keep_md: true  
---

1. The data set at [http://www.stats.uwo.ca/faculty/braun/data/rnf6080.dat] records hourly rainfall at a certain location in Canada, every day from 1960 to 1980.
    a. First, we need to load the data set into R using the command `read.table()`. Use the help function to learn what arguments this function takes. Once you have the necessary input, load the data set into R and make it a data frame called `rain.df`.
```{r}
rain <- read.table('http://www.stats.uwo.ca/faculty/braun/data/rnf6080.dat')
```
    b. How many rows and columns does `rain.df` have? (If there are not 5070 rows and 27 columns, something is wrong; check the previous part to see what might have gone wrong in the previous part.)
```{r}
nrow(rain)
ncol(rain)
```

    c. What are the names of the columns of `rain.df`?
v1-v27


    d. What is the value of row 5, column 7 of `rain.df`?
```{r}
rain[5,7]
```

    e. Display the second row of `rain.df` in its entirety.
```{r}
rain[2,]

```

    f. Explain what this command does:
```{r]}
names(rain) <- c("year","month","day",seq(0,23))
```
by running it on your data and examining the object. (You may find the display functions `head()` and `tail()` useful here.) Is it clear now what the last 24 columns represent?

//names the columns. last 24 are the hour

    g. Create a new column in the data frame called `daily`, which is the sum of the rightmost 24 columns. With this column, create a histogram of the values in this column, which are supposed to be daily rainfall values. What is wrong with this picture?  
```{r}
rain$daily = apply(rain[,4:27], 1, sum)
hist(rain$daily)
```
//negative rail fall?


h. Create a new data frame `rain.df.fixed` that takes the original and fixes it for the apparent flaw you have discovered. Having done this, produce a new histogram with the corrected data and explain why this is more reasonable.
```{r}
rain[rain < 0] <- 0
rain$daily = apply(rain[,4:27], 1, sum)
hist(rain$daily)
```


//Throws out negitive values

2. Syntax and class-typing.
    a. For each of the following commands, either explain why they should be errors, or explain the non-erroneous result.

```
vector1 <- c("5", "12", "7", "32")
max(vector1)
sort(vector1)
sum(vector1)
```
//can't sum characters

    b. For the next series of commands, either explain their results, or why they should produce errors.

```
vector2 <- c("5",7,12)
vector2[2] + vector2[3]
#numbers get cast as characters, which can't be added

dataframe3 <- data.frame(z1="5",z2=7,z3=12)
dataframe3[1,2] + dataframe3[1,3]
#19 - dataframes can have different types

list4 <- list(z1="6", z2=42, z3="49", z4=126)
list4[[2]]+list4[[4]]  #168 - lists can have different types
list4[2]+list4[4]  #single brackets give back a list which can't be added
```

3. Working with functions and operators.
    a. The colon operator will create a sequence of integers in order. It is a special case of the function `seq()` which you saw earlier in this assignment. Using the help command `?seq` to learn about the function, design an expression that will give you the sequence of numbers from 1 to 10000 in increments of 372.
```{r}
    seq(from=1, to=10000, by=372)
    ```
    
Design another that will give you a sequence between 1 and 10000 that is exactly 50 numbers in length.
```{r}
seq(from=1, to=10000, length.out=50)
```

    b. The function `rep()` repeats a vector some number of times. Explain the difference between `rep(1:3, times=3) and rep(1:3, each=3).
    
    #times repeats c(1, 2, 3) 3 times, each 1 1 1 2 2 2 3 3 3