---
   title: "hw3"  
   author: "Adam Pearce"  
   output:  
     html_document:  
       keep_md: true  

---
```{r setup, include = FALSE, cache = FALSE}
knitr::opts_chunk$set(error = TRUE)
```

Load data from URL
```{r, echo=FALSE}
library(ggplot2)
library(dplyr)

gdURL  <-  "http://tiny.cc/gapminder"
gDat  <- read.delim(file = gdURL)

gtbl <- tbl_df(gDat)
```

###Get the maximum and minimum of GDP per capita for all continents.

```{r}
GDPextent <- gtbl %>%
  select(year, country, continent, gdpPercap) %>%
  arrange(continent, year) %>%
  group_by(continent) %>%
  filter(min_rank(desc(gdpPercap)) < 2 | min_rank(gdpPercap) < 2)

(GDPextent)
```

Scatter plot connecting max and min gdpPerCap
```{r, echo=FALSE}
ggplot(GDPextent, aes(x = year, y = gdpPercap, color=continent)) + geom_point() +geom_line()
```

Haiti has suffered a steep decline
```{r, echo=FALSE}
ggplot(gtbl  %>% filter(country == "Haiti"), aes(x = year, y = gdpPercap, color=country)) + geom_point() + geom_line()
```

###Compute a trimmed mean of life expectancy for different years. Or a weighted mean, weighting by population. Just try something other than the plain vanilla mean.
```{r}
meanByYear  <- gtbl  %>% 
  group_by(year)  %>%  
  summarise(mean = mean(lifeExp), wmean = weighted.mean(lifeExp, pop))
```


```{r}
ggplot(meanByYear, aes(x = year, y = wmean)) + geom_point() + geom_line()
```

###The determine how many countries on each continent have a life expectancy less than this benchmark, for each year.

```{r}
gtbl  <-  gtbl %>%
  mutate(bench = meanByYear$wmean[match(year, meanByYear$year)],
        benchDif = lifeExp - bench,
        belowBench = lifeExp < bench)

head(gtbl)
```

Generate grouped data
```{r}
byContYear = gtbl  %>% 
    group_by(year, continent)  %>% 
    summarize(n_obs = n(),
              percentBelow = mean(belowBench))
```

Show change over time
```{r}
ggplot(byContYear, aes(x = year, y = percentBelow, color=continent)) + geom_point() +geom_line()
```